---
title: "LilyChang_Capstone_Cleaning"
output: html_notebook
---


```{r}
library(tidyverse)
```


```{r}
dataset <- read_csv("Noise_Complaints.csv")
construction <- read_csv("New_Home_Builders_20260310.csv")
alcohol <- read_csv("Department_of_Alcohol_Beverage_Services_(ABS)_Licensee_Data_20260324.csv")

```

### Cleaning main dataset (Noise Complaints)

```{r}
names(dataset) <- gsub(" ","_", names(dataset))
names(dataset) <- gsub("-","_", names(dataset))
names(dataset) <- tolower(names(dataset))

names(dataset)[names(dataset) == "address_zip_code"] <- "zip_code"

dataset$zip_code <- ifelse(
  grepl("^\\d{5}$", dataset$zip_code),
  dataset$zip_code,
  "missing"
)
head(dataset)
```


```{r}
leaf_blower <- c("Leaf Blower Sales","Noise Industrial/commercial   Leaf Blower Noise")
leaf_blower_gas <- c("Leaf Blower Use   Petro Fueled")
residential_mechanical <- c("Residential   Mechanical")
residential_motor_vehicle <- c("Residential   Motor Vehicle Noise")
residential_construction <- c("Industrial/commercial  construction/repair/demolit", "Residential   Construction/repair/demolition")
industrial_or_comm_equip <- c("Industrial/commercial   Refuse Noise Collection", "Industrial/commercial   Equipment Noise", "Industrial/commercial   heat Pump/ac Noise")
residentual_multi_tennant_dwelling <- c("Residential Muti Tenant Dwelling")
other_causes <- c("Waiver Request", "Project   Community Noise Survey", "Information Request")
residentual_house_yard <- c("Residential House/yard")
amp_equiptment <- c("Residential   El  Amplification Equip", "Industrial/commercial   Electrical Amp Equip")

dataset_mutated <- dataset |>
  mutate(case_sub_type = case_when(
    case_sub_type %in% leaf_blower ~ "leaf_blower",
    case_sub_type %in% leaf_blower_gas ~ "leaf_blower_gas",
    case_sub_type %in% residential_mechanical ~ "residential_mechanical",
    case_sub_type %in% residential_motor_vehicle ~ "residential_motor_vehicle",
    case_sub_type %in% residential_construction ~ "residential_construction",
    case_sub_type %in% industrial_or_comm_equip ~ "industrial_or_comm_equip",
    case_sub_type %in% residentual_multi_tennant_dwelling ~ "residentual_multi_tennant_dwelling",
    case_sub_type %in% other_causes ~ "other_causes",
    case_sub_type %in% residentual_house_yard ~ "residentual_house_yard",
    case_sub_type %in% amp_equiptment ~ "amp_equiptment"

 ))
```


I replaced the NAs within the zipcode column with "missing" to prevent adding more bias by removing them.

Got the code from: https://www.statology.org/replace-na-with-string-in-r/#:~:text=by%20Zach%20Bobbitt%20Published%20on%20June%202%2C,%25%3E%25%20replace_na(list(x%20=%20'missing'%2C%20y%20=%20'none')) 
and https://stackoverflow.com/questions/45101670/how-do-i-replace-na-values-with-specific-values-in-an-r 

```{r}
dataset_mutated$zip_code[is.na(dataset_mutated$zip_code)] <- "missing"
```

Code for mutating a months variable was obtained from here: https://stackoverflow.com/questions/71413141/set-2-digits-in-month 
I am doing this so I can see how the data acts within the year/s and not just from year to year.

```{r}
dataset_clean <- dataset_mutated |>
  mutate(
    parsed_date = mdy(case_open_date),                
    months = format(parsed_date, "%m")    
  )
dataset_clean
```

#### Starting the Construction Dataset cleaning

```{r}
names(construction) <- gsub(" ","_", names(construction))
names(construction) <- gsub("-","_", names(construction))
names(construction) <- tolower(names(construction))
head(construction)
```


Taking out some columns that are NA heavy and irrelevant, though I don't really have to worry much about the NAs since I'm only going to use this dataset for mapping, and the location variables have few NAs.

I also created year variables, one from the issue date of the licence, one from the expiration date of the license. 

```{r}
construction_filtered <- construction |>
  select(!c("address_line_2", "additional_applicant", "trade_name", "business_name"))

construction_issue <- construction_filtered |>
  mutate(year_of_issue = year(mdy(issue_date)))

construction_clean <- construction_issue |>
  mutate(year_of_exp = year(mdy(expiration_date)))

construction_clean$location[is.na(construction_clean$location)] <- "missing"

construction_clean$city <- tolower(construction_clean$city)

construction_clean
```

#### Starting the Alcohol Lisenceing Dataset cleaning

I will likely use this dataset minimally. I wanted it for multiple reasons, but the main one being that a major issue for more recent years was amp_equipment, or amplification equipment. this aligns with MoCo's actions against private, for-profit parties causing a lot of noise issues. Obviously, there isn't a dataset on parties, and any about rented houses (that could've been used for the parties) weren't very good. So, I went to what is often associated with parties and loud individuals: alcohol. 

Short article on the parties: https://www.nbcwashington.com/news/local/noisy-for-profit-parties-at-private-homes-face-higher-fines-in-montgomery-co/3957191/ 

```{r}
names(alcohol) <- gsub(" ","_", names(alcohol))
names(alcohol) <- gsub("-","_", names(alcohol))
names(alcohol) <- tolower(names(alcohol))
head(alcohol)
```

Dealing with the NAs in the location variable in this dataset the same way I did previously.
```{r}
alcohol$location[is.na(alcohol$location)] <- "missing"

```


I noticed the city variable needed some work, so I used mutate and case_when to clean it up.

```{r}
unique(alcohol$city)
```


```{r}
silver_spring <- c("SILVER SPRING", "SILVER SRING", "SILVER SRPING")
bethesda <- c("BETHEDA", "BETHESDA")
ashton <- c("ASHTON")
north_bethesda <- c("NORTH BETHESDA")
boyds <- c("BOYDS")
beltsville <- c("BELTSVILLE")
glen_echo <- c("GLEN ECHO")
brookeville <- c("BROOKEVILLE", "BROOKVILLE")
wheaton <- c("WHEATON", "WEATON")
poolesville <- c("POOLESVILLE")
olney <- c("OLNEY")
gaithersburg <- c("GAITHERSURG", "GAITHERSBURG", "Gaithersburg")
unknown <- c("MD")
chevy_chase <- c("CHEVY CHASE")
potomac <- c("POTOMAC")
cabin_john <- c("CABIN JOHN")
dickerson <- c("DICKERSON")
laytonsville <- c("LAYTONSVILLE")
mount_airy <- c("MOUNT AIRY")
germantown <- c("GERMANTOWN")
damascus <- c("DAMASCUS")
kensington <- c("KENSINGTON")
derwood <- c("DERWOOD")
montgomery_village <- c("MONTGOMERY VILLAGE")
beallsville <- c("BEALLSVILLE")
colesville <- c("COLESVILLE")
sandy_spring <- c("SANDY SPRING")
rockville<- c("ROCKVILLE")
takoma_park <- c("TAKOMA PARK")
burtonsville <- c("BURTONSVILLE")
garrett_park <- c("GARRETT PARK")
clarksburg <- c("CLARKSBURG")
spencerville <- c("SPENCERVILLE")




alcohol_clean <- alcohol |>
  mutate(city = case_when(
    city %in% silver_spring ~ "silver_spring",
    city %in% bethesda ~ "bethesda",
    city %in% ashton ~ "ashton",
    city %in% north_bethesda ~ "north_bethesda",
    city %in% boyds ~ "boyds",
    city %in% beltsville ~ "beltsville",
    city %in% glen_echo ~ "glen_echo",
    city %in% brookeville ~ "brookeville",
    city %in% wheaton ~ "wheaton",
    city %in% poolesville ~ "poolesville",
    city %in% olney ~ "olney",
    city %in% gaithersburg ~ "gaithersburg",
    city %in% unknown ~ "unknown",
    city %in% chevy_chase ~ "chevy_chase",
    city %in% potomac ~ "potomac",
    city %in% cabin_john ~ "cabin_john",
    city %in% dickerson ~ "dickerson",
    city %in% laytonsville ~ "laytonsville",
    city %in% mount_airy ~ "mount_airy",
    city %in% germantown ~ "germantown",
    city %in% damascus ~ "damascus",
    city %in% kensington ~ "kensington",
    city %in% derwood ~ "derwood",
    city %in% montgomery_village ~ "montgomery_village",
    city %in% beallsville ~ "beallsville",
    city %in% colesville ~ "colesville",
    city %in% sandy_spring ~ "sandy_spring",
 city %in% rockville ~ "rockville",
     city %in% takoma_park ~ "takoma_park",
     city %in% burtonsville ~ "burtonsville",
    city %in% garrett_park ~ "garrett_park",
    city %in% clarksburg ~ "clarksburg",
    city %in% spencerville ~ "spencerville"

 ))

```

checking that it looks how I want it to.
```{r}
head(alcohol_clean)
```








