---
title: "LilyChang_EDA"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(zipcodeR)
library(sf)
library(RColorBrewer)

```


```{r}
dataset <- read_csv("noise_complaints_cleaned.csv")
construction <- read_csv("new_homes_cleaned.csv")
alcohol <- read_csv("alcohol_licensing_cleaned.csv")

```
### EDA on the Noise Complaints dataset

```{r}
bargraph1 <- ggplot(dataset, aes(case_sub_type, fill = case_sub_type)) + geom_bar() +
  labs(
    title = "Counts for Reasoning behind Complaints",
    x = "case sub type"
  ) +
theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
    ) + scale_fill_brewer(palette = "Set1")
bargraph1
```

```{r}
count_by_category <- dataset |>
  count(case_year)

ggplot(count_by_category, aes(x = case_year, y = n, group = 1)) +
  geom_line() +
  labs(
    y = "Count",
    x = "Case Year",
    title = "Noise Complaints 1989-2026") 
```

```{r}
count_zip <- dataset |>
  select(zip_code) |>
  count(zip_code) |>
  arrange(desc(n))
count_zip
```

#### These bar graphs look at 2025 and 2018 specifically

Looking at the spike in 2025 and the years leading up to it
```{r}
filteredgraph25 <- dataset |>
  select(c("case_year", "case_sub_type", "zip_code")) |>
filter(case_year %in% c("2023","2024","2025"))
```

```{r}
filteredbar25 <- ggplot(filteredgraph25, aes(case_sub_type, fill=case_sub_type)) +
  geom_bar() +
  facet_wrap(~case_year) +
theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
    ) + 
  labs(
    x = "Sub Type of the Complaint",
    title = "Counts for the case sub types from 2023 to 2025"
  ) + scale_fill_brewer(palette = "Set1")
filteredbar25
```

Looking at 2018 and the years before and after it
```{r}
filteredgraph18 <- dataset |>
  select(c("case_year", "case_sub_type", "zip_code")) |>
filter(case_year %in% c("2017","2018","2019", "2020", "2021"))
```

```{r}
filteredbar18 <- ggplot(filteredgraph18, aes(case_sub_type, fill=case_sub_type)) +
  geom_bar() + 
scale_fill_brewer(palette = "Set1") +
  facet_wrap(~case_year) +
theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
    ) + 
  labs(
    x = "Sub Type of the Complaint",
    title = "Counts for the case sub types from 2016 to 2020"
  ) 

filteredbar18
```

### Suface-level EDA on Construction and Alcohol Datasets

```{r}
count_by_categoryC <- construction |>
  count(year_of_issue)

linegraph_con <- ggplot(count_by_categoryC, aes(year_of_issue, n)) + geom_line() +
  labs(
    title = "Licenses to build new homes from 1990 to 2026",
    x = "Year the license was issued",
    y = "counts"
  )
linegraph_con
```


Finding the city with the most stores licensed to sell alcohol
```{r}
most_concentrated <- alcohol |>
  count(zip) |>
  arrange(desc(n))

most_concentrated 
```


Looking at how different complaint causes behave over time.

```{r}
count_by_category_line <- dataset |>
  count(case_sub_type, case_year) |>
  filter(!case_sub_type %in% c("other_causes"))
count_by_category_line

ggplot(count_by_category_line, aes(x = case_year, y = n, color = case_sub_type, group = case_sub_type)) +
  geom_line(linewidth = 0.9) + 
scale_color_brewer(palette = "Set1") +
  labs(
    title = "Complaint Causes Over Time",
    x = "Year, 1989-2026",
    y = "Complaint Count"
  )# + theme_dark() 
#Set3, Paired
```

### Creating my first choropleth for the noise complaints dataset
#### This is the choropleth section >:D

To complete the choropleths, I'm following the instructions from here: https://data.montgomerycountymd.gov/Technology/Geographic-data-Zip-Codes-Shape-File-/vz4m-d8ee 

geojson file obtained from dataMontgomery: https://data.montgomerycountymd.gov/Technology/Geographic-data-Zip-Codes-Shape-File-/vz4m-d8ee 

```{r}
moco_spatial <- read_sf("ZIPCODES_20260330.geojson")
ggplot(moco_spatial) +
  geom_sf(fill = "white", color = "black", linewidth = 0.3) +
  theme_void()
```
```{r}
count_zip <- dataset |>
  select(zip_code) |>
  count(zip_code) |>
  arrange(desc(n))
count_zip
```

```{r}
my_sf_merged <- moco_spatial %>%
  left_join(count_zip, by = c("zipcode" = "zip_code"))
```


```{r}
ggplot(my_sf_merged) +
  geom_sf(aes(fill = n)) +
  theme_void() +
  labs(
    title = "Zip Codes Colored by Noise Complaint Counts",
fill = "Complaint Counts"
  ) +
  geom_sf_text(aes(label = zipcode), size = 3, check_overlap = TRUE, color = "white") +
  theme_dark()
```

### Making choropleths with my construction dataset


Making sure zip code is a character. 
```{r}
construction$zip_code <- as.character(construction$zip_code)
class(construction$zip_code)

```
Notably, 20814 (Bethesda) is the top zip code for both noise complaints and new homes construction. 

```{r}
 count_zipC <- construction |>
  select(zip_code) |>
  count(zip_code) |>
  arrange(desc(n))
count_zipC
```

```{r}
my_sf_mergedC <- moco_spatial %>%
  left_join(count_zipC, by = c("zipcode" = "zip_code"))

```


The construction and noise complaint choropleths overlap heavily. 

```{r}
ggplot(my_sf_mergedC) +
  geom_sf(aes(fill = n)) +
  theme_void() +
  labs(
    title = "Zip Codes Colored by New Home Construction",
fill = "Construction License Counts"
  ) +
  geom_sf_text(aes(label = zipcode), size = 3, check_overlap = TRUE, color = "white") +
  theme_dark() 
```


Repeating the process and creating a choropleth using the same steps as before for the alcohol dataset.

Making sure zip code is a character. 
```{r}
alcohol$zip <- as.character(alcohol$zip)
class(alcohol$zip)

```
Once again, 20814 is near the top. 20910 (silver spring) is also notable zip code in both this dataset and noise complaints.

```{r}
count_zipA <- alcohol |>
  select(zip) |>
  count(zip) |>
  arrange(desc(n))
count_zipA
```

```{r}
my_sf_mergedA <- moco_spatial %>%
  left_join(count_zipA, by = c("zipcode" = "zip"))

```

```{r}
ggplot(my_sf_mergedA) +
  geom_sf(aes(fill = n)) +
  theme_void() +
  labs(
    title = "Zip Codes Colored by Locations that Sell Alcohol",
fill = "Licensed Company Counts"
  ) +
  geom_sf_text(aes(label = zipcode), size = 3, check_overlap = TRUE, color = "white") +
  theme_dark() 
```




