---
title: "LilyChang_models"
output: html_notebook
---

loading the libraries and the dataset
```{r}
library(tidyverse)
library(forecast)
library(tseries)
```


```{r}

uncleaned_dataset <- read_csv("Noise_Complaints.csv")
dataset <- read_csv("noise_complaints_cleaned.csv")
construction <- read_csv("New_Home_Builders_20260310.csv")
alcohol <- read_csv("Department_of_Alcohol_Beverage_Services_(ABS)_Licensee_Data_20260324.csv")

```

yoinked the code for fixing the zipcode variable from here: https://stackoverflow.com/questions/1435435/javascript-regular-expression-to-match-5-or-9-digit-zipcode 

```{r}
names(uncleaned_dataset) <- gsub(" ","_", names(uncleaned_dataset))
names(uncleaned_dataset) <- gsub("-","_", names(uncleaned_dataset))
names(uncleaned_dataset) <- tolower(names(uncleaned_dataset))

names(uncleaned_dataset)[names(uncleaned_dataset) == "address_zip_code"] <- "zip_code"

dataset$zip_code <- ifelse(
  grepl("^\\d{5}$", dataset$zip_code),
  dataset$zip_code,
  "missing"
)
```


```{r}
leaf_blower <- c("Leaf Blower Sales","Noise Industrial/commercial   Leaf Blower Noise")
leaf_blower_gas <- c("Leaf Blower Use   Petro Fueled")
residential_mechanical <- c("Residential   Mechanical")
residential_motor_vehicle <- c("Residential   Motor Vehicle Noise")
residential_construction <- c("Industrial/commercial  construction/repair/demolit", "Residential   Construction/repair/demolition")
industrial_or_comm_equip <- c("Industrial/commercial   Refuse Noise Collection", "Industrial/commercial   Equipment Noise", "Industrial/commercial   heat Pump/ac Noise")
residentual_multi_tennant_dwelling <- c("Residential Muti Tenant Dwelling")
other_causes <- c("Waiver Request", "Project   Community Noise Survey", "Information Request")
residentual_house_yard <- c("Residential House/yard")
amp_equiptment <- c("Residential   El  Amplification Equip", "Industrial/commercial   Electrical Amp Equip")

dataset_mutated_gas <- uncleaned_dataset |>
  mutate(case_sub_type = case_when(
    case_sub_type %in% leaf_blower ~ "leaf_blower",
    case_sub_type %in% leaf_blower_gas ~ "leaf_blower_gas",
    case_sub_type %in% residential_mechanical ~ "residential_mechanical",
    case_sub_type %in% residential_motor_vehicle ~ "residential_motor_vehicle",
    case_sub_type %in% residential_construction ~ "residential_construction",
    case_sub_type %in% industrial_or_comm_equip ~ "industrial_or_comm_equip",
    case_sub_type %in% residentual_multi_tennant_dwelling ~ "residentual_multi_tennant_dwelling",
    case_sub_type %in% other_causes ~ "other_causes",
    case_sub_type %in% residentual_house_yard ~ "residentual_house_yard",
    case_sub_type %in% amp_equiptment ~ "amp_equiptment"

 ))
```

Enter this in the console to get the time series used in for the ARIMA model: 

yearts=ts(only_coyear$n,start=min(only_coyear$case_year), frequency = 1)


Filtering to exclude the parts of my data that disrupt the potential pattern as well as creating a dataframe that only has the year and counts.
The main disruptions that are excluded are likely due to a lack of state government funding, which likely caused sharp dips in the data due to under staffing.

```{r}
only_coyear <- dataset |>
  filter(!case_year %in% c("2026", "1994", "2002", "1989", "1990", "1991", "1992", "1993", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2025")) |>
  count(case_year) 
only_coyear
```


```{r}
plot(yearts)
```
Testing for stationary data. It is not stationary.
```{r}
adf.test(yearts)
```

#### Using diff() on my time series since I got above a 0.05 in the adf test.
```{r}
stationaryts <- diff(yearts, differences = 1)
```


```{r}
adf.test(stationaryts)

```
Logging and diffing it to see if that works better.
```{r}
loggedts <- log(yearts)
stationaryts2 <- diff(loggedts, differences = 1)
```

After both logging the time series and using diff(), the pvalue is below 0.05 and the plot looks pretty good. A little worried I might have over fit the data.

```{r}
adf.test(stationaryts2)
```

```{r}
model = auto.arima(stationaryts2,ic="aic",trace=TRUE)

```

```{r}
myforecast = forecast(model, level = c(95), h=10)
```

```{r}
plot(myforecast)
```

```{r}
acf(stationaryts2)
pacf(stationaryts2)
```

# Negative Binomial Regression model

```{r}
library(MASS)

df_counts <- dataset_mutated_gas |>
  group_by(zip_code, case_year, case_sub_type) |> 
  filter(case_year != "2026") |>
  summarise(count = n(), .groups = "drop")

options(scipen = 999)
```

This model addresses how complaint counts behave when considering both the year and the case_sub_type variables but the year is not centered.

Additionally, the init.theta is below the 10-20 range, it's still considered over dispersed. So, this model was the correct choice over Poisson.
```{r}
#model <- glm.nb(
#  count ~ case_year + factor(zip_code),
#  data = df_counts
#)
#model

model1 <- glm.nb(
  count ~ case_year + factor(zip_code) + factor(case_sub_type),
  data = df_counts
)
model1
```


This model considers the year, case_sub_type, and zip_code variables, but the year is centered.

When entered into a calculator for e^0.034537, the year coefficient becomes 1.035140328 Then convert to a percent (1.035140328 x 100), 103.5140328 percent. 100-103.5140328 = a 3.5140328 percent increase early?

```{r}
df_counts$year_centered <- df_counts$case_year - mean(df_counts$case_year)

model_exp <- glm.nb(count ~ year_centered * factor(case_sub_type) + factor(zip_code), data = df_counts)
model_exp
```
Main Takeaways:
-noise complaints are increasing slightly over time, not dramatically (excluding 2026, as it's incomplete).
- Certain zip codes experience more complaints than others.
- Complaints differ across the different complaint types.

References

https://youtu.be/lUbtXI3XdPE?si=YmB2RdJ25i7422p8 



https://dplyr.tidyverse.org/reference/summarise.html?utm_source=chatgpt.com 
