---
title: "LilyChang_arima_model"
output: html_notebook
---

loading the libraries and the dataset
```{r}
library(tidyverse)
library(forecast)
library(tseries)
```


```{r}
dataset <- read_csv("noise_complaints_cleaned.csv")

```

Enter this in the console to get the time series used in for this model: 

yearts=ts(only_coyear$n,start=min(only_coyear$case_year), frequency = 1)


Filtering to exclude the parts of my data that disrupt the potential pattern as well as creating a dataframe that only has the year and counts. 
```{r}
only_coyear <- dataset |>
  filter(!case_year %in% c("2026", "1994", "2002", "1989", "1990", "1991", "1992", "1993", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2025")) |>
  count(case_year) 
only_coyear
```


```{r}
plot(yearts)
```
Testing for stationary data. It is not stationary.
```{r}
adf.test(yearts)
```

#### Using diff() on my time series since I got above a 0.05 in the adf test.
```{r}
stationaryts <- diff(yearts, differences = 1)
```


```{r}
adf.test(stationaryts)

```
Logging and diffing it to see if that works better.
```{r}
loggedts <- log(yearts)
stationaryts2 <- diff(loggedts, differences = 1)
```

After both logging the time series and using diff(), the pvalue is below 0.05 and the plot looks pretty good. A little worried I might have over fit the data.

```{r}
adf.test(stationaryts2)
```

```{r}
model = auto.arima(stationaryts2,ic="aic",trace=TRUE)

```

```{r}
myforecast = forecast(model, level = c(95), h=10)
```

```{r}
plot(myforecast)
```

```{r}
acf(stationaryts2)
pacf(stationaryts2)
```









