---
title: "Analysis_notebook"
author: "Lily C"
date: "2026-04-28"
output: html_document
---

```{r}
library(tidyverse)
library(forecast)
library(tseries)
```

```{r}
dataset <- read_csv("noise_complaints_cleaned.csv")
construction <- read_csv("New_Home_Builders_20260310.csv")
alcohol <- read_csv("Department_of_Alcohol_Beverage_Services_(ABS)_Licensee_Data_20260324.csv")

```

## ARIMA cleaning 

Filtering to exclude the parts of my data that disrupt the potential pattern as well as creating a dataframe that only has the year and counts.
The main disruptions that are excluded are likely due to a lack of state government funding, which likely caused sharp dips in the data due to under staffing. 2026 is incomplete, so I will exclude it.
```{r}
only_coyear <- dataset |>
  filter(!case_year %in% c("2026", "1994", "2002", "1989", "1990", "1991", "1992", "1993", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2025")) |>
  count(case_year) 
only_coyear
```

## Negative Binomial Regression cleaning

Creating a counts dataframe.

'scipen' was used to keep things readable for me.
```{r}
df_counts <- dataset |>
  group_by(zip_code, case_year, case_sub_type) |> 
  filter(case_year != "2026") |>
  summarise(count = n(), .groups = "drop")

options(scipen = 999)
```



