---
title: "LilyChang_NegativeBinomialRegression"
author: "Lily C"
date: "2026-04-27"
output: html_document
---

loading the libraries and the dataset
```{r}
library(tidyverse)
library(forecast)
library(tseries)
```

```{r}
dataset <- read_csv("noise_complaints_cleaned.csv")
construction <- read_csv("New_Home_Builders_20260310.csv")
alcohol <- read_csv("Department_of_Alcohol_Beverage_Services_(ABS)_Licensee_Data_20260324.csv")

```

# Negative Binomial Regression model

```{r}
library(MASS)

df_counts <- dataset |>
  group_by(zip_code, case_year, case_sub_type) |> 
  filter(case_year != "2026") |>
  summarise(count = n(), .groups = "drop")

options(scipen = 999)
```

This model addresses how complaint counts behave when considering both the year and the case_sub_type variables but the year is not centered.

Additionally, the init.theta is below the 10-20 range, it's considered over dispersed. So, this model was the correct choice over Poisson.
```{r}

model <- glm.nb(
  count ~ case_year + factor(zip_code) + factor(case_sub_type),
  data = df_counts
)
model
```

This model considers the year, case_sub_type, and zip_code variables, but the year is centered.

When entered into a calculator for e^0.0283741 , the year coefficient becomes 1.028780479 Then convert to a percent (1.028780479 x 100), 102.8780479 percent. 100-102.8780479 = a 2.878047922 percent increase early.

```{r}
df_counts$year_centered <- df_counts$case_year - mean(df_counts$case_year)

model_cen <- glm.nb(count ~ year_centered * factor(case_sub_type) + factor(zip_code), data = df_counts)
model_cen
```

Main Takeaways:
-noise complaints are increasing slightly over time, not dramatically (excluding 2026, as it's incomplete).
- Certain zip codes experience more complaints than others.
- Complaints differ across the different complaint types.

